Supplementary Information

### 2020-12-21 doc/manuscript/MEE_manuscript.Rmd
update results

## Comparison with traditional methods {-#Comparison}

Using $RW$ as the growth metric, our process picked out similar climate drivers to what would be obtained via traditional methods (Figs. `r SCBI_example`, `r all_traditional_comparisons`; Table `r qualitative_results_comparison`; Appendix `r traditional_comparison`).
While one-to-one correspondence of estimated climate coefficients was neither expected nor observed, estimates were correlated and rarely differed significantly (Appendix `r traditional_comparison`; `r all_traditional_comparisons`).

@@

## Climate variable selection {-#Climate}

```{r , eval=FALSE, echo=FALSE}
## Tallies for the following paragraph (Dec.21)
#sameV_sameW: LDWp, NMt, NIOt, NIOp, SCBIp, SCt, ZOFp
#sameV_diffW: BCIp, BCIt, CBp, CBt, HFt, HKKp, LDWt, NMp, SCBIt, SCp, ZOFt
#diffV_diffW: HFp, HKKt
```
The three metrics of growth ($RW$, $BAI$, and $\Delta AGB$), consistently exhibited similar strength of correlation and direction of response to climate variables across the range of potential time windows, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).
In some cases (n= 7 of 20), both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_sameV_sameW`). 
In 11 cases, *climwin* identified the same climate variable but different (often overlapping) time windows (e.g., Fig. `r climwin_sameV_diffW`). 
Finally, two cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at HKK, precipitation variable group at HF; Figs. `r clim_sens`, `r GLS_HKK`, `r GLS_HF`), *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_diffV_diffW`). 
Henceforth, we focus on the climate drivers identified when $RW$ was the growth metric and for the full set of cores (*i.e.*, including those for which $DBH$ could not be reconstructed.)

```{r , eval=FALSE, echo=FALSE}
## Tallies for the above paragraph (Dec.21)
#PPT over PDF: 8 sites with both. PDF selected in HKK, HF, ZOF, Niobrara
#T group:Tmax at 5, PET at 3, Tmin at 1, Tmp at 1
```
Within the precipitation group, precipitation amount ($PPT$) and frequency ($PDF$) were each identified as the top variable in four of the eight sites with both variables available (Fig. `r clim_sens`). 
At the one site for which streamflow data were available (NIO), $PDF$ was selected as a stronger--*albiet* weak--driver (Figs. `r clim_sens`).
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, which were identified by *climwin* as the top temperature-related driver at five and three of the ten sites, respectively, noting that $PET$ was not available for two of these sites. 
$T_{mean}$ and $T_{min}$ were each identified as the top driver at only one site (NIO and BCI, respectively), noting that the effects of $T_{min}$ at BCI were only marginally significant (Fig. `r clim_sens`).

@@

>>>>

Using $RW$ as the growth metric, our process identified similar climate sensitivities to those identified via traditional analysis methods (Figs. `r SCBI_example`, `r all_traditional_comparisons`; Table `r qualitative_results_comparison`; Appendix `r traditional_comparison`).
While one-to-one correspondence of estimated climate coefficients was neither expected nor observed, estimates were correlated and rarely differed significantly (Appendix `r traditional_comparison`; `r all_traditional_comparisons`).

@@

```{r , eval=FALSE, echo=FALSE}
## Tallies for the following paragraph (Dec.21)
#sameV_sameW: LDWp, NMt, NIOt, NIOp, SCBIp, SCt, ZOFp
#sameV_diffW: BCIp, BCIt, CBp, CBt, HFt, HKKp, LDWt, NMp, SCBIt, SCp, ZOFt
#diffV_diffW: HFp, HKKt
```
The three metrics of growth ($RW$, $BAI$, and $\Delta AGB$)  exhibited similar strength of correlation and direction of response to climate variables across the range of potential time windows, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).
In 7 of 20 cases, both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_sameV_sameW`). 
In 11 cases, *climwin* identified the same climate variable but different (often overlapping) time windows (e.g., Fig. `r climwin_sameV_diffW`). 
Finally, two cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at HKK, precipitation variable group at HF; Figs. `r clim_sens`, `r GLS_HKK`, `r GLS_HF`), *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_diffV_diffW`). 
Henceforth, unless otherwise noted, we focus on the climate drivers identified when $RW$ was the growth metric and for the full set of cores (*i.e.*, including those for which $DBH$ could not be reconstructed.)

```{r , eval=FALSE, echo=FALSE}
## Tallies for the above paragraph (Dec.21)
#PPT over PDF: 8 sites with both. PDF selected in HKK, HF, ZOF, Niobrara
#T group:Tmax at 5, PET at 3, Tmin at 1, Tmp at 1
```
Within the precipitation group, precipitation amount ($PPT$) and frequency ($PDF$) were each identified as the top variable in four of the eight sites with both variables available (Fig. `r clim_sens`). 
At the one site for which streamflow data were available (NIO), $PDF$ was selected as a stronger--*albiet* weak--driver (Figs. `r clim_sens`).

The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, which were identified by *climwin* as the top temperature-related driver at five and three of the ten sites, respectively, noting that $PET$ was not available for two of these sites. 
$T_{mean}$ and $T_{min}$ were each identified as the top driver at only one site (NIO and BCI, respectively), noting that the effects of $T_{min}$ at BCI were only marginally significant (Fig. `r clim_sens`).

@@

(*Results up to this point are current as of Dec. 21. Those following need to be updated. I also want to reorganize to integrate with the content above*)
